Predicting Correct Dumbbell Lifts Using Data From Personal Fitness Devices
==========================================================================

####We examined accelerometer data gathered from personal fitness devices such as Jawbone Up, Nike FuelBand and Fitbit to see if we could predict whether users performing dumbell lifts were doing so with correct form.

```{r}
library(rpart)
library(caret)
library(party)
```

###Read the training and testing data and do a bit of cleaning

```{r}
setwd("~/Desktop/Fitness Device Project - Machine Learning")

# replace DIV/0 errors with NA

training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))

# remove columns X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window

training <- training[colnames(training[-(1:7)])]

testing <- testing[colnames(testing[-(1:7)])]

# make all formats numeric except for last column (classe)

for(i in c(1:152)) {training[,i] = as.numeric(as.character(training[,i]))}
for(i in c(1:152)) {testing[,i] = as.numeric(as.character(testing[,i]))}

# remove near zero value columns in both sets

nzv_vars <- nzv(training, saveMetrics = TRUE)

remove_list_nzv_vars <- rownames(nzv_vars[which(nzv_vars$nzv==TRUE),])

# create training and testing sets without nzv_vars

remove_list <- names(training) %in% remove_list_nzv_vars

training <- training[!remove_list]
testing <- testing[!remove_list]

```

###Partition large training set into 80% sub-training set and 20% sub-testing set

```{r}

# create testing and training subsets

inTrain <- createDataPartition(y=training$classe, p=0.8, list=FALSE )
sub_training <- training[inTrain,]
sub_testing <- training[-inTrain,]

```

###Create model from training set

```{r}
# create model

set.seed(793462)

fit <- rpart(classe ~ ., data=sub_training, method="class")

predictions <- predict(fit, sub_testing, type = "class")

```

###Look at the confusion matrix to check accuracy

```{r}
# look at the confusion matrix

confusionMatrix(predictions, sub_testing$classe)

```

### Now use on testing set
```{r}
predictions_final <- predict(fit, testing, type="class")

```

### For better accuracy, let's try using random forest

```{r}


data.controls <- cforest_unbiased(ntree=200, mtry=3, trace=F)

data.cforest <- cforest(classe ~., data = sub_training, controls=data.controls)

sub_testing.pred <- predict(data.cforest, sub_testing, OOB=TRUE)

confusionMatrix(table(sub_testing.pred, sub_testing$classe))

# make final prediction for 20 row set

predict(data.cforest, testing, OOB=TRUE)


```

